Hierarchical Time Series Forecasting in Emergency Medical Services

### 2023-11-10 manuscript/main.tex
Ran styler to obtain better code formatting

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{todonotes,mathtools,bm,amsmath,mathpazo}
\mathtoolsset{showonlyrefs}
\setlength{\parindent}{0cm}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{Journal of Service Research}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Hierarchical Time Series Forecasting in Emergency Medical Services},
  pdfkeywords={healthcare, emergency services, forecast
reconciliation, ambulance demand, regression},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\begin{document}

@@

\begin{keyword}
    healthcare \sep emergency services \sep forecast
reconciliation \sep ambulance demand \sep 
    regression
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, sharp corners, breakable, enhanced]}{\end{tcolorbox}}\fi

\hypertarget{sec-intro}{%
\section{Introduction}\label{sec-intro}}

@@

However, despite a large number of studies dedicated to forecasting for
EMS
\citep{mingliterature2022, gul2020exhaustive, ibrahim2016modeling, wargon2009systematic},
the hierarchical data structure has been largely ignored, and the main
focus has been on producing independent (base) forecasts at a single
level. Generating independent forecasts can result in a lack of
consistency and coordination, and therefore leads to less effective
planning and decision making.
\textcolor{teal}{For example, suppose the annual budget for the whole organization is to be allocated to each area health board based on demand within the corresponding area. If the forecasts are incoherent, there is a mismatch between the total resources to be allocated, and the resources allocated to each area. Further, suppose the budget also needs to take account of the nature of incidents that occur within each area, with more money allocated for some types of incidents than others. Then we need forecasts of the demand disaggregated by nature of incidents and health board, but these data are often too noisy to forecast directly. Forecast reconciliation addresses these problems by ensuring the forecasts are coherent across all aggregated and disaggregated series (avoiding mismatches), and by using the signal in the aggregated data to allow forecasting of highly disaggregated data (allowing better targetting of the budget).}

@@

\hypertarget{sec-lit}{%
\section{Research background}\label{sec-lit}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Current studies ignore the inherent hierarchical and/or grouped
  structure of the time series data, and the relationship between series
  at different levels of hierarchy. While the hierarchical forecasting
  methodology has been developed and applied in various domains over the
  past 10 years \citep{panagiotelis2022probabilistic}, it has never been
  explored in this area.
\item
  Current research is mainly concerned with generating point forecasts
  at a single level of hierarchy. There is a lack of studies considering
  the entire forecast distribution of daily ambulance demand for the
  whole hierarchy to better represent the uncertainty of future demand,
  providing a risk management tool for planners.
\item
  Reproducibility is still a major challenge in EMS forecasting, as it
  is unlikely that any reader can reproduce prior studies without the
  help of the authors of those papers.
\item
  Another limitation is related to the generated forecasts not being on
  the sample space of non-negative counts. Since actual ambulance counts
  cannot be negative or non-integer, ambulance demand forecast
  distributions should reflect the data. Of course, point forecasts
  represent means, so they should be non-negative, but may be
  non-integer. While this might not be an issue when producing forecasts
  at a single level, producing non-negative count forecasts in a
  hierarchical/grouped structure is challenging and requires further
  investigation in the future.
\end{enumerate}

@@

\hypertarget{tbl-literature}{}
\begin{landscape}\begin{table}[!h]
\caption{\label{tbl-literature}Summary of some studies on forecasting in Emergency Medical Services }\tabularnewline

@@

\resizebox{\linewidth}{!}{
\begin{tabular}{lr>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{4em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{8em}ll}
\toprule
Reference & Year & Variable & Horizon & Method & Metric & Probabilistic & Reconciliation\\
\midrule
Current study & 2023 & Ambulance demand & 84 days & Stationary,  Exponential Smoothing State Space (ETS), Poisson regression using Generalized Linear Model (GLM) and tscount (TSGLM), a simple empirical distribution and an ensemble method & MASE, MSSE, CRPS & YES & YES\\
Al-Azzani et al. & 2021 & Call volume & 7 , 30 ,  90 days & ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis & RMSE, MAPE & NO & NO\\
Haugsbø et al. & 2021 & Ambulance demand in Spatio-temporal & 1hour & MLP, LSTM & MAE, MSE, Cross-Entropy & YES & NO\\
Grekousis et al. & 2019 & Locations of incidents & 1 hour & MLP and Hungarian algorithm & RMSE & NO & NO\\
Ibrahim et al. & 2016 & Call volume & 1, 7 , 14 days & multiplicative univariate forecasting, univariate mixed-effects, bivariate mixed-effects model, and bivariate mixed-effects & RMSE, prediction interval coverage & Partial & NO\\
Vile et al. & 2012 & Ambulance demand & 7, 14, 21,  28 day & Singular Spectrum Analysis, ARIMA,  Holt-Winters & RMSE & NO & NO\\
Sasaki et al. & 2010 & Ambulance demand & 5 years & OLS regression & R\textasciicircum{}2 & NO & NO\\
\bottomrule
\end{tabular}}
\endgroup{}
\end{table}
\end{landscape}

\hypertarget{sec-experiment}{%
\section{Experiment setup}\label{sec-experiment}}

@@

\hypertarget{sec-data}{%
\subsection{Data}\label{sec-data}}

@@

{\centering

\raisebox{-\height}{

\includegraphics{main_files/figure-pdf/fig-hierarchy-1.pdf}

@@

@@

\subcaption{\label{fig-hierarchy-1}Hierarchical structure: Attended
incidents in the whole country are disaggregated into 3 control areas
and then into 7 different healthboards.}
\end{minipage}%
%
\begin{minipage}[t]{0.02\linewidth}

{\centering

~

@@

\end{minipage}%
%
\begin{minipage}[t]{0.42\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{../img/group.png}

\subcaption{\label{fig-hierarchy-2}Grouped structure: Incidents could be
grouped into priority (i.e.~Red, Amber \& Green) and the nature of
attended incident (i.e.~there are 35 different nature of incidents
including chest pain, breathing problems, heart attack, stroke, and so
on). The symbol * refers to the crossed attributes between hierarchical
and grouped levels.}
\end{minipage}%

\end{figure}

\hypertarget{tbl-hierarchy}{}
\begin{table}
\caption{\label{tbl-hierarchy}Number of time series in each level for the hierarchical \& grouped
structure of attended incidents }\tabularnewline

\centering
\begin{threeparttable}
\begin{tabular}{lr}
\toprule
Level & Number of series\\
\midrule
All country & 1\\
Control & 3\\
Health board & 7\\
Priority & 3\\
Priority * Control & 9\\
\addlinespace
Priority * Health board & 21\\
Nature of incident & 35\\
Nature of incident * Control & 105\\
Nature of incident * Health board & 245\\
Priority * Nature of incident & 104\\
\addlinespace
Control * Priority * Nature of incident & 306\\
Control * Health board * Priority * Nature of incident (Bottom level) & 691\\
Total & 1530\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item \textit{Note: } 
\item Due to certain combinations of the nature of incident with other variables, there is a lack of representation in the dataset. As a result, for example, instead of the calculation 3 * 35 = 105, it would be modified to 3 * 35-1 = 104.
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

{\centering \includegraphics[width=0.7\textwidth,height=\textheight]{main_files/figure-pdf/fig-feature-1.pdf}

@@

\end{figure}

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-dataviz2-1.pdf}

@@

\hypertarget{forecasting-methods}{%
\subsection{Forecasting methods}\label{forecasting-methods}}

@@

\hypertarget{performance-evaluation}{%
\subsection{Performance evaluation}\label{performance-evaluation}}

@@

A related measure is MSSE
\citep{hyndman2021forecasting, makridakis2022m5}, which uses squared
errors rather than absolute errors:
\begin{equation}\protect\hypertarget{eq-RMSSE}{}{
  \text{MSSE} = \text{mean}(q_{j}^2),
}\label{eq-RMSSE}\end{equation} where, \[
  q^2_{j} = \frac{ e^2_{j}}
 {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T (y_{t}-y_{t-m})^2},
\] Again, this is scale-independent, and smaller MSSE values suggest
more accurate forecasts.

@@

To measure the forecast distribution accuracy, we calculate the
Continuous Rank Probability Score
\citep{gneiting2014probabilistic, hyndman2021forecasting}. It rewards
sharpness and penalizes miscalibration, so it measures overall
performance of the forecast distribution.
\begin{equation}\protect\hypertarget{eq-CRPS}{}{
  \text{CRPS} = \text{mean}(p_j),
}\label{eq-CRPS}\end{equation} where \[
  p_j = \int_{-\infty}^{\infty} \left(G_j(x) - F_j(x)\right)^2dx,
\] where \(G_j(x)\) is the forecasted probability distribution function
for forecast horizon \(j\), and \(F_j(x)\) is the true probability
distribution function for the same period.

@@

\hypertarget{sec-htc}{%
\section{Hierarchical and grouped time series forecasting
techniques}\label{sec-htc}}

@@

\hypertarget{independent-base-forecast}{%
\subsection{Independent (base
forecast)}\label{independent-base-forecast}}

@@

\hypertarget{reconciliation-methods}{%
\subsection{Reconciliation methods}\label{reconciliation-methods}}

@@

\hypertarget{notations}{%
\subsubsection{Notations}\label{notations}}

@@

\hypertarget{bottom-up-bu-and-linear-reconciliation-methods}{%
\subsubsection{Bottom-up (BU) and linear reconciliation
methods}\label{bottom-up-bu-and-linear-reconciliation-methods}}

@@

Linear reconciliation methods can be written \citep{WicEtAl2019} as \[
  \tilde{\bm{y}}_h = \bm{S}(\bm{S}'\bm{W}^{-1}\bm{S})^{-1}\bm{W}^{-1}\hat{\bm{y}}_h,
\] where \(\bm{W}\) is an \(n \times n\) positive definite matrix, and
\(\hat{\bm{y}}_h\) contains the \(h\)-step forecasts of \(\bm{y}_{T+h}\)
given data to time \(T\). Different choices for \(\bm{W}\) lead to
different solutions such as Ordinary Least Squares (OLS), Weighted Least
Squares (WLS) and Minimum Trace (MinT).

\textcolor{teal}{Ordinary Least Squares (OLS) is the simplest and most commonly used method for estimating the parameters in linear regression models. In this approach, the estimation of $\bm{W}$ is based on the assumption that all the errors have equal variance. Hence, $\bm{W}$ is simply defined as the identity matrix multiplied by a constant factor. The intuition behind OLS is that it minimizes the sum of squared residuals between the observed and predicted values of the dependent variable. The main weakness of this approach is that it does not take account of the different scales of the base time series; the aggregated series will usually have higher variance than the disaggregated series, simply because the values are larger, but OLS treats all series the same. A strength of the approach is that it is simple, and does not involve estimating a covariance matrix.}

\textcolor{teal}{Weighted Least Squares (WLS) is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each series. But it assumes that the errors of each series are uncorrelated with each other. In this approach, $\bm{W}$ is defined as a diagonal matrix with the variance of the errors on the diagonal. The intuition behind WLS is that it assigns higher weight to series with smaller error variance, and thereby takes into account the different scales of the base time series. The main weakness of this approach is that it ignores the relationships between series. A strength of WLS is that it is relatively easy to compute $\bm{W}$ as it is based only on error variances which are readily estimated.}

\textcolor{teal}{Minimum Trace (MinT) is a further generalization where $\bm{W}$ is defined as the covariance matrix of the base forecast errors. So it takes account of both the scale of each series, and the relationships between the series.}
\citet{WicEtAl2019}
\textcolor{teal}{showed that this approach gives the optimal reconciled forecasts in the sense that the sum of the forecast variances is minimized. The main weakness of this approach is that it is difficult to estimate the full covariance matrix. In practice, we usually need to use a shrinkage estimate where the off-diagonal elements are shrunk towards zero.}

@@

\hypertarget{sec-results}{%
\section{Results and discussion}\label{sec-results}}

@@

The overall forecasting performance is reported in
Table~\ref{tbl-result}, in which the average forecast accuracy over
horizons 43--84 days (corresponding to the planning horizon) is
presented per model, method, and the hierarchical level. Reported
forecast accuracy is averaged across all forecast horizons, rolling
origins, and series at each level. Table~\ref{tbl-result} presents both
point and probabilistic forecast accuracy at total, control area, health
board and bottom-level series. Point forecast performance is reported
using MASE and MSSE in Table~\ref{tbl-result-1} and \ref{tbl-result-2},
respectively. Probabilistic forecast accuracy is reported using CRPS in
Table~\ref{tbl-result-3}. The bold entries in each table identify a
combination of method and model that performs best for the corresponding
level (i.e.~each column), based on the smallest values of accuracy
measures.

Table~\ref{tbl-result-1} and \ref{tbl-result-2} show that forecast
reconciliation (i.e.~MinT) improves forecast accuracy at the higher
levels of the hierarchy including total, control area and health board.
However, it does not result in accuracy improvement at the bottom-level
series, for which base forecasts are more accurate. This might be due to
the noisy structure of time series at the bottom level, and possibly due
to very different patterns in the aggregated series. It is also clear
from Table~\ref{tbl-result-1} that the ensemble method improves forecast
accuracy at total, control area and health board. However, this does not
remain valid for bottom series where different individual methods
perform best, depending on the accuracy measure.
\textcolor{teal}{While the forecast reconciliation approach aims to enhance forecast accuracy, its effectiveness is not guaranteed, especially if the bottom-level series exhibit excessive noise and lack systematic patterns. Despite this, reconciling forecasts at the bottom level can offer advantages by generating coherent forecasts that facilitate alignment in planning across various teams within an organization, promote better coordination, and prevent conflicting decisions. Moreover, even when dealing with noisy and irregular bottom-level series, reconciliation can still improve forecast accuracy at higher levels of the hierarchy by leveraging the information available across the hierarchy. Therefore, although the bottom-level forecasts may not be highly accurate on their own, reconciling them with higher-level forecasts can still provide a more consistent view of future demand and potentially yield more accurate forecasts at other levels.}

@@

\caption{\label{tbl-result}Average forecast performance calculated on
the test sets at forecast horizons \(h=43,\dots,84\) days, with time
series cross validation applied to attended incident data. The testset
consists of 462 days. The best approach is highlighted in
bold.}\begin{minipage}[t]{\linewidth}
\subcaption{\label{tbl-result-1}Point forecast accuracy using MASE }

{\centering

\tabularnewline

>>>>

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{todonotes,mathtools,bm,amsmath,mathpazo}
\mathtoolsset{showonlyrefs}
\setlength{\parindent}{0cm}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Journal of Service Research}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Hierarchical Time Series Forecasting in Emergency Medical Services},
  pdfkeywords={healthcare, emergency services, forecast
reconciliation, ambulance demand, regression},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

@@

\begin{keyword}
    healthcare \sep emergency services \sep forecast
reconciliation \sep ambulance demand \sep 
    regression
\end{keyword}
\end{frontmatter}
    \section{Introduction}\label{sec-intro}

@@

To illustrate the problem, let's consider a very simple example where we
have an EMS provider with a national level of governance, and two
regions (A and B), each with a health board and a station. There is a
total national budget to be split between the regions in proportion to
the forecast number of incidents in each region. The two regions have
very different incident patterns, and so must be forecast using
different models. However, the data are noisy at regional level, so the
national forecasts are best obtained by summing the demand from the two
regions. The resulting national forecasts are not equal to the sum of
the regional forecasts, and so are not coherent. In fact, the national
forecasts show a decreasing trend in demand, and so the national
governing body decides to cut the budget for the next year. But neither
of the regional forecasts shows a trend, and so the regions argue that
the budget cut is unfair. In addition, Region A has much more variable
demand than Region B, and so to cope with periods of peak demand, Region
A needs to hold more resources in reserve. So the budget distribution
needs to be made in a way that ensures the probability of each region
being unable to meet demand is equal. Our solution to this problem is to
use a hierarchical forecasting approach that ensures the forecasts are
probabilistically coherent. Then any trends or other forecast
characteristics at national level will also be reflected in the regional
forecasts, and the probabilistic forecasts allow for the different
levels of uncertainty in the two regions. Budget can be allocated by
controlling the probability of demand exceeding available resources,
rather than being simply in proportion to the expected demand.

Despite a large number of studies dedicated to forecasting for EMS
\citep{mingliterature2022, gul2020exhaustive, ibrahim2016modeling, wargon2009systematic},
the hierarchical data structure has been largely ignored, and the main
focus has been on producing independent (base) forecasts at a single
level. Generating independent forecasts can result in a lack of
consistency and coordination, and therefore leads to less effective
planning and decision making.
\textcolor{teal}{For example, suppose the annual budget for the whole organization is to be allocated to each area health board based on demand within the corresponding area. If the forecasts are incoherent, there is a mismatch between the total resources to be allocated, and the resources allocated to each area. Further, suppose the budget also needs to take account of the nature of incidents that occur within each area, with more money allocated for some types of incidents than others. Then we need forecasts of the demand disaggregated by nature of incidents and health board, but these data are often too noisy to forecast directly. Forecast reconciliation addresses these problems by ensuring the forecasts are coherent across all aggregated and disaggregated series (avoiding mismatches), and by using the signal in the aggregated data to allow forecasting of highly disaggregated data (allowing better targetting of the budget).}

@@

\section{Research background}\label{sec-lit}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Current studies ignore the inherent hierarchical and/or grouped
  structure of the time series data, and the relationship between series
  at different levels of hierarchy. While the hierarchical forecasting
  methodology has been developed and applied in various domains over the
  past 10 years \citep{panagiotelis2023probabilistic}, it has never been
  explored in this area.
\item
  Current research is mainly concerned with generating point forecasts
  at a single level of hierarchy. There is a lack of studies considering
  the entire forecast distribution of daily ambulance demand for the
  whole hierarchy to better represent the uncertainty of future demand,
  providing a risk management tool for planners.
\item
  Reproducibility is still a major challenge in EMS forecasting, as it
  is unlikely that any reader can reproduce prior studies without the
  help of the authors of those papers.
\item
  Another limitation is related to the generated forecasts not being on
  the sample space of non-negative counts. Since actual ambulance counts
  cannot be negative or non-integer, ambulance demand forecast
  distributions should reflect the data. Of course, point forecasts
  represent means, so they should be non-negative, but may be
  non-integer. While this might not be an issue when producing forecasts
  at a single level, producing non-negative count forecasts in a
  hierarchical/grouped structure is challenging and requires further
  investigation in the future.
\end{enumerate}

@@

\caption{\label{tbl-literature}Summary of some studies on forecasting in
Emergency Medical Services}

\centering{

@@

\resizebox{\linewidth}{!}{
\begin{tabular}{lr>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{4em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{8em}ll}
\toprule
Reference & Year & Variable & Horizon & Method & Metric & Probabilistic & Reconciliation\\
\midrule
Current study & 2023 & Ambulance demand & 84 days & Stationary,  Exponential Smoothing State Space (ETS), Poisson regression using Generalized Linear Model (GLM) and tscount (TSGLM), a simple empirical distribution and an ensemble method & MASE, MSSE, CRPS & YES & YES\\
Al-Azzani et al. & 2021 & Call volume & 7 , 30 ,  90 days & ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis & RMSE, MAPE & NO & NO\\
Haugsbø et al. & 2021 & Ambulance demand in Spatio-temporal & 1hour & MLP, LSTM & MAE, MSE, Cross-Entropy & YES & NO\\
Grekousis et al. & 2019 & Locations of incidents & 1 hour & MLP and Hungarian algorithm & RMSE & NO & NO\\
Ibrahim et al. & 2016 & Call volume & 1, 7 , 14 days & multiplicative univariate forecasting, univariate mixed-effects, bivariate mixed-effects model, and bivariate mixed-effects & RMSE, prediction interval coverage & Partial & NO\\
Vile et al. & 2012 & Ambulance demand & 7, 14, 21,  28 day & Singular Spectrum Analysis, ARIMA,  Holt-Winters & RMSE & NO & NO\\
Sasaki et al. & 2010 & Ambulance demand & 5 years & OLS regression & R\textasciicircum{}2 & NO & NO\\
\bottomrule
\end{tabular}}
\endgroup{}

\end{table}%

\section{Experiment setup}\label{sec-experiment}

@@

\subsection{Data}\label{sec-data}

@@

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{main_files/figure-pdf/fig-hierarchy-1.pdf}

@@

\subcaption{\label{fig-hierarchy-1}Hierarchical structure: Attended
incidents in the whole country are disaggregated into 3 control areas
and then into 7 different healthboards.}

\end{minipage}%
%
\begin{minipage}[t]{0.02\linewidth}
~\end{minipage}%
%
\begin{minipage}[t]{0.42\linewidth}

\centering{

\includegraphics[width=0.6\textwidth,height=\textheight]{../img/group.png}

@@

\subcaption{\label{fig-hierarchy-2}Grouped structure: Incidents could be
grouped into priority (i.e.~Red, Amber \& Green) and the nature of
attended incident (i.e.~there are 35 different nature of incidents
including chest pain, breathing problems, heart attack, stroke, and so
on). The symbol * refers to the crossed attributes between hierarchical
and grouped levels.}

\end{figure}%

\caption{\label{tbl-hierarchy}Number of time series in each level for
the hierarchical \& grouped structure of attended incidents}

\centering{

\centering
\begin{threeparttable}
\begin{tabular}{lr}
\toprule
Level & Number of series\\
\midrule
All country & 1\\
Control & 3\\
Health board & 7\\
Priority & 3\\
Priority * Control & 9\\
\addlinespace
Priority * Health board & 21\\
Nature of incident & 35\\
Nature of incident * Control & 105\\
Nature of incident * Health board & 245\\
Priority * Nature of incident & 104\\
\addlinespace
Control * Priority * Nature of incident & 306\\
Control * Health board * Priority * Nature of incident (Bottom level) & 691\\
Total & 1530\\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item \textit{Note: } 
\item Due to certain combinations of the nature of incident with other variables, there is a lack of representation in the dataset. As a result, for example, instead of the calculation 3 * 35 = 105, it would be modified to 3 * 35-1 = 104.
\end{tablenotes}
\end{threeparttable}

@@

\end{table}%

@@

\centering{

\includegraphics[width=0.7\textwidth,height=\textheight]{main_files/figure-pdf/fig-feature-1.pdf}

@@

\end{figure}%

@@

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-dataviz2-1.pdf}

@@

\subsection{Forecasting methods}\label{forecasting-methods}

@@

\subsection{Performance evaluation}\label{performance-evaluation}

@@

A related measure is MSSE
\citep{hyndman2021forecasting, makridakis2022m5}, which uses squared
errors rather than absolute errors:
\begin{equation}\phantomsection\label{eq-RMSSE}{
  \text{MSSE} = \text{mean}(q_{j}^2),
}\end{equation} where, \[
  q^2_{j} = \frac{ e^2_{j}}
 {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T (y_{t}-y_{t-m})^2},
\] Again, this is scale-independent, and smaller MSSE values suggest
more accurate forecasts.

@@

To measure the forecast distribution accuracy, we calculate the
Continuous Rank Probability Score
\citep{gneiting2014probabilistic, hyndman2021forecasting}. It rewards
sharpness and penalizes miscalibration, so it measures overall
performance of the forecast distribution.
\begin{equation}\phantomsection\label{eq-CRPS}{
  \text{CRPS} = \text{mean}(p_j),
}\end{equation} where \[
  p_j = \int_{-\infty}^{\infty} \left(G_j(x) - F_j(x)\right)^2dx,
\] where \(G_j(x)\) is the forecasted probability distribution function
for forecast horizon \(j\), and \(F_j(x)\) is the true probability
distribution function for the same period.

@@

\section{Hierarchical and grouped time series forecasting
techniques}\label{sec-htc}

@@

\subsection{Independent (base
forecast)}\label{independent-base-forecast}

@@

\subsection{Reconciliation methods}\label{reconciliation-methods}

@@

\subsubsection{Notations}\label{notations}

@@

\subsubsection{Bottom-up (BU) and linear reconciliation
methods}\label{bottom-up-bu-and-linear-reconciliation-methods}

@@

Linear forecast reconciliation methods can be written
\citep{WicEtAl2019} as \[
  \tilde{\bm{y}}_h = \bm{S}(\bm{S}'\bm{W}^{-1}\bm{S})^{-1}\bm{W}^{-1}\hat{\bm{y}}_h,
\] where \(\bm{W}\) is an \(n \times n\) positive definite matrix, and
\(\hat{\bm{y}}_h\) contains the \(h\)-step forecasts of \(\bm{y}_{T+h}\)
given data to time \(T\). When \(\bm{W}_h\) is the covariance matrix of
\(\hat{\bm{y}}_h\), the resulting forecasts are optimal in the sense
that the sum of the variances of the reconciled forecasts is minimized,
provided the base forecasts \(\hat{\bm{y}}_h\) are unbiased. However,
\(\bm{W}_h\) is difficult to estimate, and so there have been various
suggested approximations to \(\bm{W}_h\), leading to different types of
reconciliation such as Ordinary Least Squares (OLS), Weighted Least
Squares (WLS) and Minimum Trace (MinT).

\textcolor{teal}{Ordinary Least Squares (OLS) is the simplest and most commonly used method. In this approach, the estimation of $\bm{W}$ is based on the assumption that all the errors are uncorrelated and have equal variance, and that multi-step forecast variances are proportion to one-step forecast variances. Then, $\bm{W}$ is simply the identity matrix multiplied by a constant factor. The main weakness of this approach is that it does not take account of the different scales of the base time series; the aggregated series will usually have higher variance than the disaggregated series, simply because the values are larger, but OLS treats all series the same. A strength of the approach is that it is simple, and does not involve estimating a covariance matrix.}

\textcolor{teal}{Weighted Least Squares (WLS) is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each series. But it assumes that the errors of each series are uncorrelated with each other, and that multi-step forecast variances are proportion to one-step forecast variances. In this approach, $\bm{W}$ is defined as a diagonal matrix with the variance of the errors on the diagonal. The intuition behind WLS is that it assigns higher weight to series with smaller error variance, and thereby takes into account the different scales of the base time series. The main weakness of this approach is that it ignores the relationships between series. A strength of WLS is that it is relatively easy to compute $\bm{W}$ as it is based only on error variances which are readily estimated.}

\textcolor{teal}{Minimum Trace (MinT) is a further generalization where $\bm{W}$ is defined as the covariance matrix of the one-step base forecast errors. So it takes account of both the scale of each series, and the relationships between the series.}
But it still assumes that multi-step forecast variances are proportion
to one-step forecast variances. The main weakness of this approach is
that it is difficult to estimate the full covariance matrix, even of the
one-step errors. In practice, we usually need to use a shrinkage
estimate where the off-diagonal elements are shrunk towards zero.\}

@@

\section{Results and discussion}\label{sec-results}

@@

The overall forecasting performance is reported in
Table~\ref{tbl-result}, in which the average forecast accuracy over
horizons 43--84 days (corresponding to the planning horizon) is
presented per model, method, and the hierarchical level. Reported
forecast accuracy is averaged across all forecast horizons, rolling
origins, and series at each level. Table~\ref{tbl-result} presents both
point and probabilistic forecast accuracy at total, control area, health
board and bottom-level series. Point forecast performance is reported
using MASE and MSSE, while probabilistic forecast accuracy is reported
using CRPS. The bold entries in each table identify a combination of
method and model that performs best for the corresponding level
(i.e.~each column), based on the smallest values of accuracy measures.

Table~\ref{tbl-result} shows that forecast reconciliation (i.e.~MinT)
improves point forecast accuracy at the higher levels of the hierarchy
including total, control area and health board. However, it does not
result in accuracy improvement at the bottom-level series, for which
base forecasts are more accurate. This might be due to the noisy
structure of time series at the bottom level, and possibly due to very
different patterns in the aggregated series. It is also clear from
Table~\ref{tbl-result} that the ensemble method improves forecast
accuracy at total, control area and health board. However, this does not
remain valid for bottom series where different individual methods
perform best, depending on the accuracy measure.
\textcolor{teal}{While the forecast reconciliation approach aims to enhance forecast accuracy, its effectiveness is not guaranteed, especially if the bottom-level series exhibit excessive noise and lack systematic patterns. Despite this, reconciling forecasts at the bottom level can offer advantages by generating coherent forecasts that facilitate alignment in planning across various teams within an organization, promote better coordination, and prevent conflicting decisions. Moreover, even when dealing with noisy and irregular bottom-level series, reconciliation can still improve forecast accuracy at higher levels of the hierarchy by leveraging the information available across the hierarchy. Therefore, although the bottom-level forecasts may not be highly accurate on their own, reconciling them with higher-level forecasts can still provide a more consistent view of future demand and potentially yield more accurate forecasts at other levels.}

@@

\caption{\label{tbl-result}Average forecast performance calculated on
the test sets at forecast horizons \(h=43,\dots,84\) days, with time
series cross validation applied to attended incident data. The test set
consists of 462 days. The best approach is highlighted in bold. Point
forecast accuracy is measured using MASE and MSSE, while probabilistic
forecast accuracy is measured using CRPS.}

\centering{
